Ice-shelf ocean boundary layer dynamics from large-eddy simulations

### 2021-10-06 Response_reviewers_rev1.tex

\documentclass[tc, manuscript]{copernicus}

\setcounter{figure}{0}
\begin{document}

\title{Response to reviewers for ``Ice-shelf ocean boundary layer dynamics from large-eddy simulations"}

\runningtitle{IOBL dynamics}

\runningauthor{Begeman et al.}

\correspondence{Carolyn Begeman (cbegeman@lanl.gov)}

\Author[1]{Carolyn Branecky}{Begeman}
\Author[1]{Xylar}{Asay-Davis}
\Author[1]{Luke}{Van Roekel}

\affil[1]{Los Alamos National Laboratory, P.O. Box 1663, Los Alamos, New Mexico, USA 87545}

\firstpage{1}

\maketitle

\section{Response to Reviewer 1}

Abstract:

Line 2: “Yet these small scale processes, which regulate heat transfer between….”\\
Should be heat and salt transfer.\\
\textit{The suggested change has been made.}

Introduction:

Line 27: “….an overturning circulation known as “ice pump” “
Could talk about the refreezing process.\\
\textit{We have added the following sentence to this paragraph:}
``Supercooling of the IOBL and frazil ice accretion to the ice-shelf base are also regionally important processes for cold cavities, but are not the focus of this work \citep{galton-fenzi_modeling_2012, jordan_conditional_2015}.''

Line 31: “ IOBLs present unique conditions in the global ocean, involving a stabilizing flux from phase change” The stabilizing flux mostly comes from the freshening of the boundary layer not from ‘phase-change’.\\
\textit{We intended to specifically refer to the stabilizing buoyancy flux at the boundary associated with ice-shelf melting. We hope this revision is more clear:}\\
``IOBLs present unique conditions in the global ocean, involving a stabilizing buoyancy flux from melting ice and a boundary layer that is positively buoyant against a sloping boundary.''

Line 36: “ Numerical studies addressing intermediate scale”
Please define ‘Intermediate scale’. Also to my knowledge Mondal et al., 2019 was a DNS study that resolved the Kolmogorov scale.\\
\textit{We have avoided this vague term in our revision. The revised sentence reads:}\\
``IOBL turbulence has been explored through laboratory experiments, direct numerical simulations, and large-eddy simulations \citep{middleton_numerical_2021, mondal_ablation_2019, vreugdenhil_stratification_2019, mcconnochie_dissolution_2018}.''

should clearly mention the author used a Lewis number =1.\\
\textit{This comment wasn't entirely clear to us. If the reviewer is referring to this study, it isn't the case that the Lewis number=1 either in the melt parameterization or the AMD turbulence closure model.}

It would be interesting to see how friction length changes over slope, assuming eddy viscousity is estimated as a product of friction velocity and mixing length.\\
%XX

Methods:

General Comment: I would love a schematic of the model set up.\\
\textit{We have added a schematic figure, now labeled Figure 1.}

Overview of the LES model: It would help the readers if you clearly mention whether you have used a linear or non-linear equation of state.\\
\textit{Thanks for catching the omission. We have added the following text:}\\
``We use the nonlinear equation of state from  \citet{jackett_algorithms_2006} to compute densities.''

Simulation set-up:

Line 150: What about the time step?
\textit{We have added the following text:}
``PALM employs adaptive timestepping; timesteps for the simulations presented here range from 0.5 s to 2.75 s after 2 h.''
\textit{We also realized that the following information might be useful to the reader:}
``We initiate turbulence over first 50 min of the simulation with perturbations to the horizontal velocity components on the order of 0.01 m/s.''

Line 167: I don’t think 1 degree is a relatively high slope for Antarctic ice-shelves.
\textit{That's fair. We have deleted the judgment so that it now reads:}
``and a slope of 1\unit{^{\circ}}''

Result:

A plot with temporal evolution of melt rate, obukhov lengthscales could help the reader.
\textit{We have included a plot with the temporal evolution of melt rates, Figure 1 c,f. We believe that the obukhov lengthscale doesn't provide much additional context beyond that given by the melt rate and friction velocity in Figure 1, as it is computed from them. However, we added the following text to provide some context for the reader:}
``Average melt rates over the last inertial period range from 0.3--2.5\unit\,{m} (Fig. \ref{fig:timeseries}a,d) and average Monin-Obukhov lengthscales are 3--5\,\unit{m} (not shown).''

Discussion :

Line 397: 'The relationship between melt rate and distant thermal driving' : the sentence is hard to read.
\textit{To make our point clear, we have rewritten the latter portion of this paragraph:}
``The relationship between average ice-shelf melt rates and the thermal driving for the cavity as a whole can be conceptualized in two components. The first is the relationship between the local thermal driving and melt rate, determined primarily by the local ocean turbulence. The second is the relationship between distant thermal driving (i.e., the water masses entering the ice-shelf cavity) and the strength of sub-ice-shelf circulation \citep{holland_response_2008}. Only the former is addressed by this study. Our simulations do not capture the large-scale increase in overturning circulation that accompanies an increase in distant thermal driving. Our simulations partially capture the increase in IOBL velocity due to an increase in thermal driving but the far-field velocity is unaffected. We note that the often cited quadratic relationship between distant thermal driving and melt rate involves both local turbulent processes and large-scale processes \citep{holland_response_2008}; studies generally find this exponent on distant thermal driving to be between 1.5 and 2 \citep{favier_assessment_2019, jourdain_ocean_2017, little_how_2009}.''

\section{Response to Reviewer 2}

This manuscript addresses boundary layer dynamics at the ice-ocean interface motivated by obtaining novel parameterizations of ice-shelf melting. This is a highly important topic: increasing complexity of ice and ocean models will not improve the estimates of melt rates with inappropriate representation of the processes in the boundary layer. Although the method chosen in the paper (LES using relatively low resolution) gives the results of limited applicability, which is admitted by the authors, it provides some important insights. The manuscript can be published if the authors address the following points

Major points:

Section 2.3: Please add the figure illustrating simulation setup for your base case. Add more detail in this figure: currently even the sizes of the domain are not written anywhere.

\textit{We have added a schematic figure (Figure 1). This text specifying the domain size was included in the original manuscript, but it would be easy to miss. With the new figure, it should be more visible}\\
``The domain is a 64 m$^3$ cube''

L 100: please add more detail about validation in the appendix. It would be useful to have an idea how the chosen model works compared to other model (which one? Was it DNS or measurements or another LES model?).

\textit{We have added a figure to the supplemental materials (Figure S1) and the following text to the manuscript. We don't feel that too much information is needed about model set-up because we follow Abkar and Moin (2017) exactly.}\\
``We validated our implementation against the stable atmospheric boundary layer test case published in \citet{abkar_large-eddy_2017} and \citet{stoll_large-eddy_2008}. We follow \citet[]['s]{abkar_large-eddy_2017}'s model set-up exactly and chose their moderate resolution with 72x72x72 grid points. Our results are in good agreement with \citet[]['s]{abkar_large-eddy_2017} solution employing another LES model with AMD (boundary layer height differs by <10\%) and other LES models with different turbulence closures (Fig. S1, their Fig. 1 and references therein).''

L 150: How does the resolution influence the results? Why this particular resolution was chosen?

L 157-158: It would be helpful to write the boundary conditions explicitly, not just Dirichlet/von Neumann. 
\textit{We have modified this paragraph to explicitly state the boundary conditions, and rearrange some of the text accordingly. The relevant sentences are:}
``We set von Neumann boundary conditions at the top boundary corresponding to the dynamic sub-grid momentum and scalar fluxes (Equations~\ref{eq:tau}, \ref{eq:vflux_heat}, and \ref{eq:vflux_salt}), as resolved fluxes go to zero at a no-penetration boundary. ... Boundary conditions are Dirichlet at the bottom boundary, set to the far-field temperature, salinity, and geostrophic velocity. ... The flow is periodic along the x and y dimensions.''
If there is a flow developing along the ice face due to buoyancy, how justified is the use of the periodic boundary conditions?
\textit{In the literature, periodic boundary conditions in the context of simulations of the IOBL have been employed either to represent the domain moving with the IOBL in a Lagrangian sense (Chen et al. 2021. The Cryosphere Discussions.) or the domain remaining fixed in space and the periodic boundary conditions providing the opportunity for the development of turbulence and of equilibrium conditions (Vreugdenhil and Taylor, 2019; Rosevear et al. 2021). We take the latter view, with the intent to bring the flow to equilibrium with the forcing by the end of the simulations. As discussed in Section 4.1, equilibrium was not achieved. }

The whole Sec. 2.2, called ‘Turbulence closure’, is about boundary conditions at the ice-ocean interface.
\textit{We agree that this may not be the clearest descriptor of this section. We move the first paragraph to the previous section and rename this section ``Boundary conditions for ice-shelf melting.''}

Section Results should be split into smaller subsections with separate titles to make it more reader-friendly.
\textit{Good suggestion. We have added the following subsections: Overview of the mean simulated state, Turbulent kinetic energy budget, Boundary layer turbulence, Melting and its relation to thermal driving and slope, Vertical structure of turbulent fluxes.}

Minor points:

Please check the equations. Several equations have misprints:
\textit{Thank you for reviewing the equations so carefully. We have made the changes you requested}

Eq (1), rhs: 1st term: should be $x_j$ and not $x_i$; 3rd term: what does index g in $u_g$ mean?
\textit{We have added the following text to define $u_g$:}
``The momentum terms on the right hand side of Equation \ref{eq:uprog} are, in order: advection, Coriolis forcing, imposed geostrophic flow with a geostrophic velocity $u_g$, ...''

Eq (3),(4): 2nd terms in the rhs: should be $u_j$ not $u_i$;
\textit{Change has been made.}

Eq (11): not clear, does this mean that gamma is the sum of the molecular and turbulent coefficients? 
\textit{The text has been modified to make this equation more clear:}
``where the exchange coefficients are defined with two terms, one representing turbulent transfer within the turbulent surface layer and one representing molecular transfer within the viscous sublayer following \citet{mcphee_dynamics_1987} (their Equation 11):.''
Why the form with power ‘-1’ is chosen if that power can be removed from this equation. In addition, I have not found this or similar expression in McPhee et al (1987), please clarify.
\textit{The power of -1 was not necessary in these equations an has been removed. This form (without the -1 power) can be found in Equation 11 of McPhee et al. (1987) where the second term is the molecular contribution.}

Eq (14) $c_p$ should be removed, it makes the expression dimensionally inconsistent.
\textit{Change has been made.}

L 146: ‘It is noted by the PALM developers that this error was found to be small’ –please provide a reference here.
\textit{This note is embedded in the code itself.}
\url{https://github.com/lanl/palm_lanl/blob/7f1444073b20ae81b451d4ab469d13e66b5358b7/trunk/SOURCE/surface_layer_fluxes_mod.f90#L2513-L2519}

L 168: far field thermal driving is 0.15C, but in the figures elsewhere, there is no such case, only 0.1C. I believe one of those is a misprint.

L 147: ‘This error would be small if the first ∼10 cm were largely a constant flux layer, as hypothesized by McPhee (1983).’ – I do not understand this sentence. Clearly, if 10 cm were a constant flux layer, then the mentioned fluxes would be equal. What was hypothesized by McPhee?
\textit{We have edited the text to be a bit more specific about what we mean. We hoped to convey that the surface layer is thought to deviate only slightly from a constant flux layer. We also changed the citation to McPhee (1981), which contains this argument in more detail (though the 1983 paper provides the extension to scalar fluxes).}``This error would be small if the first $\sim$10\,\unit{cm} were nearly a constant flux layer; \citet{mcphee_analytic_1981} hypothesized that the sub-ice surface layer would be nearly but not exactly a constant flux layer.''

L 205: ‘thermal driving increases from 0.5 to 0.6C ’: there is no 0.5C case in the figure.
XX

Fig. 1: the caption does not match the figure: (a) is not TKE but friction velocity etc.
\textit{We have replaced figure 1 with the version with TKE in panels a,d.}

L 347: critical gradient Richardson number for development of hydrodynamic instabilities, and further transition to turbulence, $Ri_g$=0.25, has been first obtained by Miles and Howard in 1961.
\textit{Thanks for the tip. This reference has been added.}

\section{Community comments}
I believe sections 1 and 2 would benefit greatly from a figure showing a sketch of the simulated flow, and some relevant physical quantities and conditions, e.g. the angles alpha and beta, etc.

\textit{We have added a schematic figure, now labeled Figure 1.}

Also, if I'm not mistaking, the numerical methods used by the LES solver are not discussed. They should at least be mentioned briefly.
\textit{This is a good suggestion, particularly as PALM has several methods available for advection, timestepping, and pressure. We have added the following text:}
``We employ PALM's implementation of the \citet{piacsek_conservation_1970} second-order advection scheme for momentum and scalars and a third-order Runge-Kutta timestepping scheme. We also use the \citet{temperton_generalized_1992} Fast-Fourier Transform algorithm for the pressure solver.''

\subsection{Unsolicited changes}
\textit{We have also added the following citations:}
Rosevear XX
\end{document}

### 2021-10-06 main_JGR.tex

\documentclass[tc, manuscript]{copernicus}

@@

\correspondence{Carolyn Begeman (cbegeman@lanl.gov)}

\Author[1]{Carolyn Branecky}{Begeman}
\Author[1]{Xylar}{Asay-Davis}
\Author[1]{Luke}{Van Roekel}

\affil[1]{Los Alamos National Laboratory, P.O. Box 1663, Los Alamos, New Mexico, USA 87545}

\runningtitle{IOBL dynamics}

\runningauthor{Begeman et al.}

\received{}
\pubdiscuss{}
\revised{}
\accepted{}
\published{}

\firstpage{1}

\maketitle

\begin{abstract}
Small scale, turbulent flow below ice shelves is regionally isolated and difficult to measure and simulate.  Yet these small scale processes, which regulate heat transfer between the ocean and ice shelves, can affect sea-level rise by altering the ability of Antarctic ice shelves to “buttress” ice flux to the ocean. In this study, we improve our understanding of turbulence below ice shelves by means of large-eddy simulations at sub-meter resolution, capturing boundary layer mixing at scales intermediate between laboratory experiments or direct numerical simulations and regional or global ocean circulation models. Our simulations feature the development of an ice-shelf ocean boundary layer through dynamic ice melting in a regime with low thermal driving, low ice-shelf basal slope, and strong shear driven by the geostrophic flow. 
We present a preliminary assessment of existing ice-shelf basal melt parameterizations adopted in single component or coupled ice-sheet and ocean models on the basis of a small parameter study. While the parameterized linear relationship between ice-shelf melt rate and far-field ocean temperature appears to be robust, we point out a little-considered relationship between ice-shelf basal slope and melting worthy of further study.
\end{abstract}

\introduction

% Motivation - SLR
The largest source of uncertainty in future sea level rise is the potential loss of ice from the Antarctic Ice Sheet \citep{ipcc_climate_2014}. The rate of grounded ice loss is highly sensitive to the melting of ice shelves, which drain over 80\% of Antarctica’s grounded ice \citep{reese_far_2018, rignot_ice-shelf_2013}. The Ice-shelf Ocean Boundary Layer (IOBL) controls ice-shelf melting by regulating oceanic heat and salt fluxes to the ice shelf base. Thus, accurate predictions of ice-shelf melting depend on representing the turbulent dynamics of the IOBL. This representation is also critical for evaluating the sensitivity of the coupled land ice-ocean system to changes in ocean conditions. Of particular concern is the sensitivity of ice-shelf melting to increasing seawater temperature at depth, a trend observed along a wide swath of the West Antarctic coastline and a potential trigger for West Antarctic Ice Sheet collapse \citep{purkey_unabated_2018, ruan_ice-shelf_2021, schmidtko_multidecadal_2014, wahlin_pathways_2021}.

% TODO coauthor order
One indication that ocean models do not capture turbulent dynamics in the IOBL is that the simulated thermal driving, the difference between ocean temperature and the local freezing point, and consequently the simulated melt rate differ significantly between ocean models and as resolution is varied within a model \citep{gwyther_cold_2020}. Furthermore, ocean models predict ice-shelf melting using parameterizations that neglect the effects of the lateral buoyancy gradient of the IOBL, and often of the vertical buoyancy gradient, on the efficiency of vertical mixing near the boundary. This model deficiency likely biases turbulent fluxes through the IOBL and ocean cavity circulation, which is driven in large part by the buoyant flow of water freshened by ice-shelf melting, an overturning circulation also known as the ``ice pump'' \citep{webber_impact_2018}. A new parameterization of ice-shelf melting that accounts for the dynamics of IOBL turbulence is needed to achieve a more physically based, accurate coupling of ice sheets and oceans in climate models \citep{dinniman_modeling_2016, edwards_projected_2021, gwyther_cold_2020, naughten_future_2018}.

% Overview of IOBL turbulence lit
IOBLs present unique conditions in the global ocean, involving a stabilizing flux from phase change (ice-melting) and a boundary layer that is positively buoyant against a sloping boundary. There is a rich literature on stably-stratified boundary layers (typically under constant stabilizing flux boundary conditions), but the dependence of heat, salt and momentum fluxes on stratification remains a difficult problem, especially for strongly stratified regimes \citep{zonta_stably_2018}. The IOBL has most frequently been explored at small scales through laboratory experiments and direct numerical simulations, with some recent numerical studies addressing intermediate scales \citep{middleton_numerical_2021, mondal_ablation_2019, vreugdenhil_stratification_2019}. However, this body of work has not yet matured to setting a new standard for ice-shelf melt parameterization in ocean models. Today, the most commonly used ice-shelf melt parameterization still derives from sea-ice conditions (i.e., in the absence of a slope) with some parameters tuned for ice-shelf settings \citep{holland_modeling_1999, jenkins_observation_2010, mcphee_dynamics_1987}. A knowledge gap exists in bridging IOBL dynamics across scales and characterizing the structure of buoyant plumes. Recently, this has been addressed through non-turbulence-resolving 2-d or 2.5-d models \citep{cheng_modeling_2020, jenkins_simple_2016, jenkins_shear_2021}. Notably, \citet{jenkins_shear_2021} found that vertical mixing in IOBL settings was represented poorly by the commonly-used K-Profile Parameterization (KPP) in comparison to a higher-order turbulence closure scheme, suggesting that the application of KPP in ocean models may be inappropriate in ice-shelf cavities.

% Overview of this study
In this study, we model turbulent heat, salt, and momentum fluxes through the IOBL using Large-Eddy Simulation (LES). Whereas the ocean models typically used to model sub-ice-shelf circulation lack both the resolution and appropriate parameterizations needed to capture the relevant turbulent scales for boundary layer dynamics, LES captures the dominant energy-containing scales of turbulence and represents smaller, unresolved scales with varying degrees of complexity. An effective parameterization of ice-shelf melting is likely to rest on an understanding of how turbulent mixing in the IOBL depends on stratification and shear forcing. We vary far-field ocean temperature and ice-shelf slope between model runs and characterize turbulent fluxes and ice-shelf melt rates. In this study we focus on the high-shear, low thermal driving regime. The target of previous LES studies has been on low-shear settings \citep{middleton_numerical_2021, vreugdenhil_stratification_2019}. Section~\ref{meth} describes the LES model, its turbulence closure scheme, and the set-up of our simulations. Section~\ref{results} presents the results of our simulations, with a focus on comparisons across the sampled range of thermal driving and slope. The discussion is given in two parts: Section~\ref{disc:dyn} contextualizes the features of our simulated IOBL turbulence and discusses limitations of this study, and Section~\ref{disc:prm} focuses on how this study informs parameterizations of ice-shelf melting and IOBL turbulence. We provide some closing thoughts in Section 5.

@@

The PArallelized Large eddy simulation Model (PALM) was developed at the Institute of Meteorology and Climatology at Leibniz Universitat Hannover, Germany \citep{raasch_palm-large-eddy_2001} for simulation of atmospheric and ocean boundary layers. For this application to sub-ice-shelf settings, we developed a new version based on PALM v5.0 \citep{maronga_parallelized_2015} with added features including a boundary flux scheme for the ice-ocean interface, rotating the gravity and Coriolis vectors for sloped domains, and a different turbulence closure scheme. Here, we provide a brief overview of PALM with a focus on our additions. For more details we refer readers to \citet{maronga_parallelized_2015}.

The governing equations of PALM are the following:\\
\begin{align}
    \textrm{Momentum conservation} & &
        \frac{\partial u_i}{\partial t} &= 
        -\frac{\partial u_i u_j}{\partial x_i}
        -\varepsilon_{ijk} f_j u_k 
        + \varepsilon_{i3j} f_3 u_{g,j}\nonumber\\
     &&& \quad \, + g_i \frac{\rho - \rho_a}{\rho_0} 
        - \frac{1}{\rho_0}\frac{\partial \pi^*}{\partial x_i} \nonumber\\
     &&& \quad \, - \frac{\partial}{\partial x_j}\left(\overline{u''_i u''_j} 
        - \frac{1}{3}\overline{u''_i u''_i}\delta_{ij}\right) \label{eq:uprog}\\
    \textrm{Volume conservation} & &
        \frac{\partial u_j}{\partial x_j} &= 0\label{eq:volconserv}\\
    \textrm{Heat conservation} & &
        \frac{\partial \theta}{\partial t} &= 
        -\frac{\partial u_j \theta}{\partial x_j} 
        - \frac{\partial}{\partial x_j}\left(\overline{u''_i \theta''}\right) 
        + \Phi_{\theta}\label{eq:ptprog}\\
    \textrm{Salt conservation} & &
        \frac{\partial S}{\partial t} &= 
        -\frac{\partial u_j S}{\partial x_j} 
        - \frac{\partial}{\partial x_j}\left(\overline{u''_i S''}\right) 
        + \Phi_{S}\label{eq:saprog}
\end{align}

The momentum terms on the right hand side of Equation \ref{eq:uprog} are, in order: advection, Coriolis forcing, imposed geostrophic flow, buoyancy forcing, a correction for divergence in the flow using the pressure perturbation $\pi^*$ (imposing incompressibility), and sub-grid scale momentum flux. All prognostic variables are considered filtered at the grid scale. This is typically denoted by the overbar, which we omit except for the sub-grid turbulent flux terms to emphasize that we only represent averaged effects with a turbulence closure scheme. Double primes denote sub-grid fluctuations.

We represent a sloping ice base by rotating the gravity ($\vec{g}$) and Coriolis ($\vec{f}$) vectors while keeping the domain a rectangular prism as in \citet{vreugdenhil_stratification_2019}. 
Specifically, 
\begin{align}
	\vec{g} &= g_z [\sin \alpha \sin \beta, \sin \alpha \cos \beta,\cos \alpha]\label{eq:g}\\
    \vec{f} &= 2 \Omega [\sin \phi \sin \alpha \sin \beta,\cos \phi \sin \alpha \cos \beta,\sin \phi \cos \alpha]\label{eq:f}
\end{align}
where $g_z$ is the magnitude of gravitational acceleration in the geopotential z-direction, $\beta$ is the angle the up-slope direction makes with north, and $\alpha$ is the slope angle from horizontal. For the rotated Coriolis vector $\vec{f}$, $\Omega$ is the rotation rate and $\phi$ is the latitude. Unless otherwise specified, quantities are oriented with the simulated grid.

The buoyancy term in Equation~\ref{eq:uprog} combines the contributions of along-slope pressure gradient due to the slope of the ice shelf in hydrostatic equilibrium and buoyancy due to changes in density relative to the ambient density of the water column $\rho_a$. $\rho_a$ varies along slope and with depth from the ice interface as a function of the hydrostatic pressure:
\begin{equation}
    \rho_a(x,y,z) = f_{EOS}\left(\theta^i(z),S^i(z),p(x,y,z)\right)
\end{equation}
where the superscript $i$ denotes the initial state. The reference density $\rho_0$ is evaluated in the center of the x-y plane ($\rho_0(z) = \rho_a(x_{mid},y_{mid},z)$). We neglect hydrostatic pressure gradients along slope, except through $\rho_a$ in the buoyancy term. For the maximum slope simulated in this study, the maximum hydrostatic pressure gradient is on the order of 100\,\unit{Pa\,m^{-1}}. This simplification has the advantage of avoiding pressure discontinuities across periodic boundary.

>>>>

\documentclass[draft]{styles/agujournal2019}
\usepackage{url} 
\usepackage{lineno}
\usepackage{graphicx}
\usepackage{float}
\usepackage{mathtools}
\linenumbers
\draftfalse
\journalname{JGR: Oceans}

@@

\correspondingauthor{Carolyn Begeman}{cbegeman@lanl.gov}
\authors{Carolyn Branecky Begeman\affil{1}, Xylar Asay-Davis\affil{1}, and Luke Van Roekel\affil{1}}

\affiliation{1}{Los Alamos National Laboratory, Los Alamos, New Mexico, USA}

\begin{keypoints}
\item We simulate boundary layer turbulence with prognostic ice-shelf melting
in cold cavity settings.
\item Our simulations support a linear melt response to thermal driving.
\item Buoyant mean flow can significantly enhance melt rates at ice-shelf basal slopes as low as 1 degree.
\end{keypoints}

\begin{abstract}
Small scale, turbulent flow below ice shelves is regionally isolated and difficult to measure and simulate.  Yet these small scale processes, which regulate heat transfer between the ocean and ice shelves, can affect sea-level rise by altering the ability of Antarctic ice shelves to “buttress” ice flux to the ocean. 
%Direct measurements of ocean properties below ice shelves are exceedingly rare, expensive to perform, and difficult to generalize to larger areas. Further, the few available measurements indicate that existing boundary layer theory on sub-ice-shelf turbulence (based largely on observations below sea ice) is incomplete. 
In this study, we improve our understanding of turbulence below ice shelves by means of large-eddy simulations at sub-meter resolution, capturing boundary layer mixing at scales intermediate between laboratory experiments or direct numerical simulations and regional or global ocean circulation models. 
Our simulations feature the development of an ice-ocean boundary layer through dynamic ice melting in a regime with low thermal driving, low ice-shelf slope, and strong shear driven by the geostrophic flow. 
We present a preliminary assessment of existing ice-shelf basal melt parameterizations adopted in single component or coupled ice-sheet and ocean models on the basis of a small parameter study. While the parameterized linear relationship between ice-shelf melt rate and far-field ocean temperature appears to be robust, we point out a little-considered relationship between ice-shelf basal slope and melting worthy of further study.

\textbf{Plain language summary}\\
Ice-shelf melting from below is controlled by the rate at which the ocean can transport heat vertically. Little is known about the small-scale mixing that carries this heat close to the ice base, where melting itself impedes mixing by forming a density barrier. This study uses ocean modeling at ultra-high resolution to study these mixing processes. It offers some insights into the relationship between ice-shelf melting and ocean temperature far from the ice base, which may help us project how fast ice will melt when ocean waters entering the cavity warm. This study adds to a growing body of research that indicates coarse resolution ocean models used for such projections may have to adjust how they handle ocean mixing to improve the fidelity of ice-shelf melting.

\end{abstract}
\newpage
\section{Introduction}

% Motivation - SLR
The largest source of uncertainty in future sea level rise is the potential loss of ice from the Antarctic Ice Sheet \cite{ipcc_climate_2014}. The rate of grounded ice loss is highly sensitive to the melting of ice shelves, which drain over 80\% of Antarctica’s grounded ice \cite{reese_far_2018, rignot_ice-shelf_2013}. The Ice-shelf/Ocean Boundary Layer (IOBL) controls ice-shelf melting by regulating oceanic heat and salt fluxes to the ice shelf base. Thus, accurate predictions of ice-shelf melting depend on representing the turbulent dynamics of the IOBL. This representation is also critical for evaluating the sensitivity of the coupled land ice-ocean system to changes in ocean conditions. Of particular concern is the sensitivity of ice-shelf melting to increasing seawater temperature at depth, a trend observed along a wide swath of the West Antarctic coastline and a potential trigger for West Antarctic Ice Sheet collapse \cite{ruan_ice-shelf_2021, schmidtko_multidecadal_2014, purkey_unabated_2018, wahlin_pathways_2021}.

% Shortcomings of ocean-model representations of IOBL
One indication that ocean models do not capture turbulent dynamics in the IOBL is that the simulated thermal driving, the difference between ocean temperature and the local freezing point, and consequently the simulated melt rate differ significantly between ocean models and as resolution is varied within a model \cite{gwyther_cold_2020}. Furthermore, ocean models predict ice-shelf melting using parameterizations that neglect the lateral buoyancy gradient of the IOBL and often the vertical buoyancy gradient as well, which should reduce the efficiency of vertical mixing near the boundary. This model deficiency likely biases turbulent fluxes through the IOBL and ocean cavity circulation, which is driven in large part by the buoyant flow of water freshened by ice-shelf melting, an overturning circulation also known as the ``ice pump'' \cite{webber_impact_2018}. A new parameterization of ice-shelf melting that accounts for boundary layer dynamics is needed to achieve a more physically based, accurate coupling of ice sheets and oceans in climate models \cite{gwyther_cold_2020, dinniman_modeling_2016, naughten_future_2018, edwards_projected_2021}.

% Overview of IOBL turbulence lit
IOBLs present unique conditions in the global ocean, involving a stabilizing flux from phase change (ice-melting) and a boundary layer that is positively buoyant against a sloping boundary. There is a rich literature on stably-stratified boundary layers (typically under constant stabilizing flux boundary conditions), but the dependence of heat, salt and momentum fluxes on stratification remains a difficult problem, especially for strongly stratified regimes \cite{zonta_stably_2018}. The IOBL has been most extensively explored at small scales through laboratory experiments and direct numerical simulations, with some recent numerical studies addressing intermediate scales \cite{mondal_ablation_2019, vreugdenhil_stratification_2019, middleton_numerical_2021}. However, this body of work has not yet matured to setting a new standard for ice-shelf melt parameterization in ocean models. Today, the most commonly used ice-shelf melt parameterization still derives from sea-ice conditions (i.e., in the absence of a slope) with some parameters tuned for ice-shelf settings \cite{holland_modeling_1999, jenkins_observation_2010, mcphee_dynamics_1987}. A knowledge gap exists in bridging IOBL dynamics across scales and characterizing the structure of buoyant plumes. Recently, this has been addressed through non-turbulence-resolving 2-d or 2.5-d models \cite{jenkins_simple_2016, jenkins_shear_2021, cheng_modeling_2020}. Notably, \citeA{jenkins_shear_2021} found that vertical mixing in IOBL settings was represented poorly by the commonly-used K-Profile Parameterization (KPP) in comparison to a higher-order turbulence closure scheme, suggesting that the application of KPP in ocean models may be inappropriate in ice-shelf cavities.

% Overview of this study
In this study, we model turbulent heat, salt, and momentum fluxes through the IOBL using Large-Eddy Simulation (LES). Whereas the ocean models typically used to model sub-ice-shelf circulation lack both the resolution and appropriate parameterizations needed to capture the relevant turbulent scales for boundary layer dynamics, LES captures the dominant energy-containing scales of turbulence and represents smaller, unresolved scales with varying degrees of complexity. An effective parameterization of ice-shelf melting is likely to rest on an understanding of how turbulent mixing in the IOBL depends on stratification and shear forcing. We vary far-field ocean temperature and ice-shelf slope between model runs and characterize turbulent fluxes and ice-shelf melt rates. In this study we focus on the high-shear, low thermal driving regime. The target of previous LES studies has been on low-shear settings \cite{vreugdenhil_stratification_2019, middleton_numerical_2021}. Section~\ref{meth} describes the LES model, its turbulence closure scheme, and the set-up of our simulations. Section~\ref{results} presents the results of our simulations, with a focus on comparisons across the sampled range of thermal driving and slope. The discussion is given in two parts: Section~\ref{disc:dyn} contextualizes the features of our simulated IOBL turbulence and discusses limitations of this study, and Section~\ref{disc:prm} focuses on how this study informs parameterizations of ice-shelf melting and IOBL turbulence. We provide some closing thoughts in Section \ref{conclusion}.

@@

The PArallelized Large eddy simulation Model (PALM) was developed at the Institute of Meteorology and Climatology at Leibniz Universitat Hannover, Germany \cite{raasch_palm-large-eddy_2001}. It has been applied to the simulation of atmospheric and ocean boundary layers, but it has never been used specifically to treat a sub-ice-shelf ocean boundary layer. For this application, we developed a new version based on PALM v5.0 \cite{maronga_parallelized_2015} with added features including a boundary flux scheme for the ice-ocean interface, rotating the gravity vector for sloped domains, and a different turbulence closure scheme. Here, we provide a brief overview of PALM with a focus on our additions. For more details we refer readers to \citeA{maronga_parallelized_2015}.

The governing equations of PALM are the following:\\
\begin{align}
    \textrm{Volume conservation} & &
        \frac{\partial u_j}{\partial x_j} &= 0\label{eq:volconserv}\\
    \textrm{Momentum conservation} & &
        \frac{\partial u_i}{\partial t} &= 
        -\frac{\partial u_i u_j}{\partial x_i}
        -\varepsilon_{ijk} f_j u_k 
        + \varepsilon_{i3j} f_3 u_{g,j}\nonumber\\
     &&& \quad \, + g_i \frac{\rho - \rho_a}{\rho_0} 
        - \frac{1}{\rho_0}\frac{\partial \pi^*}{\partial x_i} \nonumber\\
     &&& \quad \, - \frac{\partial}{\partial x_j}\left(\overline{u''_i u''_j} 
        - \frac{1}{3}\overline{u''_i u''_i}\delta_{ij}\right) \label{eq:uprog}\\
    \textrm{Heat conservation} & &
        \frac{\partial \theta}{\partial t} &= 
        -\frac{\partial u_j \theta}{\partial x_j} 
        - \frac{\partial}{\partial x_j}\left(\overline{u''_i \theta''}\right) 
        + \Phi_{\theta}\label{eq:ptprog}\\
    \textrm{Salt conservation} & &
        \frac{\partial S}{\partial t} &= 
        -\frac{\partial u_j S}{\partial x_j} 
        - \frac{\partial}{\partial x_j}\left(\overline{u''_i S''}\right) 
        + \Phi_{S}\label{eq:saprog}
\end{align}

The momentum terms on the right hand side of Equation \ref{eq:uprog} are, in order: advection, Coriolis forcing, imposed geostrophic flow, buoyancy forcing, a correction for divergence in the flow using the pressure perturbation $\pi^*$ (imposing incompressibility), and sub-grid scale momentum flux. Double primes denote sub-grid fluctuations. All prognostic variables are considered filtered at the grid scale. This is typically denoted by the overbar, which we omit except for the sub-grid turbulent flux terms to emphasize that we only represent averaged effects.

%\paragraph{}% Sloping surface
We represent a sloping ice base by rotating the gravity ($\textbf{g}$) and Coriolis ($\textbf{f}$) vectors while keeping the domain a rectangular prism as in \citeA{vreugdenhil_stratification_2019}. 
Specifically, 
\begin{equation} \label{eq:g}
	%\textbf{g} = g [sin \alpha,0,cos \alpha]
	\textbf{g} = g_z [\sin \alpha \sin \beta, \sin \alpha \cos \beta,\cos \alpha]
\end{equation}
\begin{equation} \label{eq:f}
    \textbf{f} = 2 \Omega [\sin \phi \sin \alpha \sin \beta,\cos \phi \sin \alpha \cos \beta,\sin \phi \cos \alpha]
\end{equation}
where $g_z$ is the magnitude of gravitational acceleration in the geopotential z-direction, $\beta$ is the angle the up-slope direction makes with north, and $\alpha$ is the slope angle from horizontal. For the rotated Coriolis parameter $f$, $\Omega$ is the rotation rate and $\phi$ is the latitude.

The buoyancy term in Equation~\ref{eq:uprog} combines the contributions of along-slope pressure gradient due to the slope of the ice shelf in hydrostatic equilibrium and buoyancy due to changes in density relative to the ambient density of the water column $\rho_a$. $\rho_a$ varies along slope and with depth from the ice interface as a function of the hydrostatic pressure:
\begin{equation}
    \rho_a(x,y,z) = f_{EOS}\left(\theta^i(z),S^i(z),p(x,y,z)\right)
\end{equation}
where the superscript $i$ denotes the initial state. The reference density $\rho_0$ is evaluated in the center of the x-y plane ($\rho_0(z) = \rho_a(x_{mid},y_{mid},z)$). We neglect hydrostatic pressure gradients along slope, except through $\rho_a$ in the buoyancy term. For the maximum slope simulated in this study, the maximum hydrostatic pressure gradient is on the order of 100\,Pa\,m\textsuperscript{-1}. This simplification has the advantage of avoiding pressure discontinuities across periodic boundary.
